Hydrogen Epoch of Reionization Array (HERA)

### 2016-04-11 manuscript.tex
workign through 5

\subsection{The ``Wedge''}
Perhaps the most important advance informing HERA's design is a
refined understanding of how smooth-spectrum foregrounds interact
with instrument chromaticity to produce a characteristic ``wedge" of
foreground leakage in Fourier space (see Fig.~\ref{fig:wedge}), 
outside of which the 21\,cm signal dominates (the ``EoR window'').  The coordinates of the wedge are discussed below.
Through theoretical and observational work
\citep{datta_etal2010,morales_et_al2012,parsons_et_al2012b,vedantham_2012,thyagarajan_et_al2013,hazelton_et_al2013,pober_etal2013b,liu_et_al2014a,liu_et_al2014b},
we have learned how the boundary between the wedge and  the EoR window is determined by the separation between antennas,
signal reflections within antennas, and the angular response of the antenna beam.  Deep integrations also show us
that, to the limits of current sensitivity, foreground emission is absent outside of the wedge; it can only 
appear there through instrumental leakage \citep{parsons_etal2014,ali_et_al2015,moore_et_al2016,kohn_et_al2016}.

@@

Another approach to improving sensitivity is to make the collecting area per element larger.  In the limit, this yields a fully filled aperture as mentioned.  A larger element limits the accessible field-of-view, which however is not a huge liability for a focussed experiment, as long as an appropriate patch of ``cold'' sky passes within its field-of-view.  One method combining both advantages is to use small elements to build up phased array ``tiles'' and then electronically produce many beams, which may get correlated.  This adds additional complexity and potential systematics, however it is the method employed by both MWA and LOFAR.  Fig. \ref{fig:boost} (right) shows the relative sensitivity of the various array relative to HERA-19, calculated for a redshift corresponding to the peak signal at a hydrogen reionization fraction of 1/2.  The circles near the array name indicate the relative sizes of the element.  Note that MWA and LOFAR are cross-hatched to indicated phased array tiles.

@@

Note that in addition to the 19 elements in South Africa (and an initial prototype near Berkeley, CA), a pair of elements at the Green Bank Observatory in West Virginia have been used extensively in the validation testing, the results of which are highlighted in the earlier referenced material and discussed below.  Also, three elements are currently under construction at the Mullard Radio Astronomyl Observatory outside Cambridge, UK for additional feed and element tests.

@@

The large collecting area of a
HERA element yields nearly 5 times the sensitivity of an MWA tile and more
than 20 times that of a PAPER element, but it does so without substantially
degrading our ability to isolate and remove foreground emission on the basis of
spectral smoothness (Fig. \ref{fig:reflectometry}, right panel).  As shown in
\Mycitet{parsons_et_al2012b}, the amplitude and timescale of signal reflections
relates directly to the leakage of smooth-spectrum foregrounds into regions of Fourier space 
used to measure reionization.
To facilitate foreground filtering, HERA's antenna element is designed to suppress reflections at long time delays.

\begin{figure}[h!]
	\begin{tabular}{ll}
	\begin{minipage}{4.2in}
\includegraphics[width=2.1in]{plots/ref_dipole_and_hera_dish.jpg}
\includegraphics[width=2.1in]{plots/orbcomm_dish_beam_map_530cm_feed.pdf}
	\end{minipage} & 
	\begin{minipage}{2.05in}
	\caption{Left: The first of three prototype dishes at NRAO--Green Bank, used for measuring beam frequency structure with reflectometry and the beam pattern at 137\,MHz by comparing satellite signals to the reference dipole in the foreground \citep{neben_et_al2016}. Right: The measured EW power pattern plotted with dashed lines marking zenith angles of 20$^\circ$, 40$^\circ$, 60$^\circ$, 80$^\circ$.} 
	\label{fig:orbcommexptandbeammap}
	\end{minipage}
	\end{tabular}
	\vspace{-15pt}
\end{figure}

\begin{figure}[h!]
	\centering
    \includegraphics[width=0.49\textwidth]{plots/s11_compare_msip.pdf}
	%\includegraphics[width=0.49\textwidth]{plots/compare_kernels_msip_subband.pdf}
\includegraphics[width=0.5\textwidth]{plots/ps1d_with_delay_kernel.pdf}
	\vspace{-25pt}
	\caption{Left: Electromagnetic simulations (black; \citealt{ewallwice_et_al2016}) of the response of the HERA dish as a function of time delay, $\tau$,
       agree with field measurements (grey; \citealt{patra_et_al2016}). 
Right: Modeled foreground power spectra with (green) and without (blue) the antenna chromaticism over a 10\,MHz subband centered at $z=7.5$. The spectral smoothness of the dish enables foreground isolation below the level of a fiducial reionization signal (black) for most $k_{\parallel}$ modes, yielding percent-level constraints on model parameters.}
	\label{fig:reflectometry}
	\vspace{-10pt}
\end{figure}

As shown in Fig.~\ref{fig:reflectometry}, HERA's element meets this requirement in
reflectometry tests, performing comparably to the PAPER element.  The
relatively short (4.5\,m) focal height of the parabolic dish 
minimizes the timescale of feed-dish reflections while maintaining the feed's
illumination of the dish.  HERA's feed design is based on the PAPER sleeved
dipole element, inverted above a backplane, with a skirting cylinder to
improve symmetry in the polarized beam response patterns.  Figure
\ref{fig:orbcommexptandbeammap} shows the beam pattern measured at 137 MHz with a beam mapping
system using the ORBCOMM satellite network \Mycitep{neben_et_al2016}.  Results
indicate an effective per-element collecting area of 93\,m$^2$, compared with the
theoretical maximum Airy response of 155 m$^2$.  The measured primary beam is
consistent with simulations at the 0.1\% to 0.5\% level, with a full width at
half maximum of $\sim$$10^\circ$, and a first sidelobe at -20 dB \citep{ewallwice_et_al2016,neben_et_al2016,patra_et_al2016,thyagarajan_et_al2016}.
Feed-to-feed coupling between adjacent antennas has been measured to be below -50 dB, indicating
that mutual coupling will not be problematic.

\subsubsection{Element Construction}
To minimize cost and aid in construction in a remote location, the construction materials are readily available standard construction materials, making for a very cost-effective design.  The feed is supported from three utility poles using line and hardware primarily from boating activities which can handle the load and external environment.  Given the close-packed design, each pole (except for the perimeter) are shared between three antennas.  This both reduces the number of poles, but also makes for a balanced load.  Note that the perimeter poles can be stayed with guy lines if needed.   The rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.  Two of the three posts are shared with adjacent antennas.

The center of the antenna is a cast concrete donut-shaped hub with 12 sleeves for PVC spars which provide the support for the surface and 12 horizontal spars which allow for additional support.  The twelve surface spars are made of 60mm PVC pipe, which are stressed along three support points to approximate a parabolic shape.  Note that an ideal moment-loaded beam actually attains a true parabola.  Note that the three support points (one at the hub, one about 3 meters out radially and one at the rim-end) are at the correct height and angle for the underlying parabola.  The PVC essentially acts as a spatial filter.

The shape between the spars is actually a one-dimensional parabola rather than a two-dimensional paraboloid, so the dish surface may more properly be called a ``facetted parabola''.  If only twelve spars were used for the entire surface, simple Ruze losses would be quite high (about 10\% at 150 MHz).  Therefore, at about 2.4 meters another horizontal member is placed, which launches another parabolic spar.  The simple Ruze loss is then less than 1\%, which is much more acceptable.  Fig \ref{fig:elementplan} shows this surface scheme and the panelization (left) and the offsets (in mm) of that scheme from an ideal parabola.  The panel that is ``left out'' is the location of a door to allow access into the hub (via a small removable door and bridge).  The additional intermediate spars also make the panel sizes much more manageable and better secured.  The panel arrangement is also set by a standard width of mesh roll of 1.22m.

\begin{figure}[h!]
\centerline{
\includegraphics[height=6cm]{plots/antPlan.png}
\includegraphics[height=6cm]{plots/offset.png} 
}
\caption{\small Element.
\label{fig:elementplan}}
\end{figure}

\subsubsection{Feed}
The feed is to be deployed along the two separate phases.  Initially, the feed reuses the dipole portion of the PAPER element, however in a new optimized cylindrical feed-cage configuration.  The optimization was based on main beam efficiency, cross-coupling (integrated feed pattern on the dish), standing waves (blockage size), and frequency response.  The original PAPER sleeved dipole is held on a PVC mast within the cylindrical structure. The design parameters were size of backplane, height of mast, and height of cylinder.  A range of conical structures were looked at but were ruled out since it was found they introduce additional frequency structure with no real improvement in performance.  The phase 1 design is documented in a memo \cite{feedmemo}, which are currently under test on the 19 elements in South Africa. Additional studies are on-going at NRAO and Cambridge University.

@@

The low noise amplifier/balun integrates directly to the base of the feed dipole in the same configuration as per PAPER (REF).  The 150m length of 75 $\Omega$ cable used in the initial design also runs down centrally and then over to the post-amplifier module, which is housed near the RFI-shielded container containing the digital electronics.  This will be replaced by the 35m cable/node design.

@@

\noindent HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments 
to cover the $uv$-plane with sub-element sampling density (see Fig.~\ref{fig:arrayConfig}). The dense core maximizes 
sensitivity on the short baselines for which PAPER's proven foreground filtering strategy works best.
The core is supplemented by 30 additional outrigger elements 
to tile the $uv$-plane with instantaneously complete sub-aperture 
sampling out to 250$\lambda$ and complete aperture-scale sampling out to 350$\lambda$ (at 150\,MHz). This sampling 
strategy suppresses grating lobes in the synthesized beam and provides information for calibrating and correcting 
direction-dependent antenna responses \citep{dillon_parsons2016}.
Even with the sub-aperture dithering and long baselines, all 350 HERA elements can be robustly calibrated by taking advantage of 
its highly redundant configuration \citep{liu_et_al2010,zheng_et_al2014}. 
Resulting calibration errors range from $\sim5\%$ (in the core) to $\sim10\%$ (for outriggers)
of the residual fractional noise per antenna after averaging.  DILLON REF

>>>>

\subsection{The ``Wedge''}
\label{sec:wedge}
Perhaps the most important advance informing HERA's design is a
refined understanding of how smooth-spectrum foregrounds interact
with instrument chromaticity to produce a characteristic ``wedge" of
foreground leakage in Fourier space (see Fig.~\ref{fig:wedge}), 
outside of which the 21\,cm signal dominates (the ``EoR window'').  The coordinates of the wedge are discussed below.
Through theoretical and observational work
\citep{datta_etal2010,morales_et_al2012,parsons_et_al2012b,vedantham_2012,thyagarajan_et_al2013,hazelton_et_al2013,pober_etal2013b,liu_et_al2014a,liu_et_al2014b},
we have learned how the boundary between the wedge and  the EoR window is determined by the separation between antennas,
signal reflections within antennas, and the angular response of the antenna beam.  Deep integrations also show us
that, to the limits of current sensitivity, foreground emission is absent outside of the wedge; it can only 
appear there through instrumental leakage \citep{parsons_etal2014,ali_et_al2015,moore_et_al2016,kohn_et_al2016}.

@@

Another approach to improving sensitivity is to make the collecting area per element larger.  In the limit, this yields a fully filled aperture as mentioned.  A larger element limits the accessible field-of-view, which however is not a huge liability for a focussed experiment, as long as an appropriate patch of ``cold'' sky passes within its field-of-view.  One method combining both advantages is to use small elements to build up phased array ``tiles'' and then electronically produce many beams, which may get correlated.  This adds additional complexity and potential systematics, however it is the method employed by both MWA and LOFAR.  Fig. \ref{fig:boost} (right) shows the relative sensitivity of the various array relative to HERA-19, calculated for a redshift corresponding to the peak signal at a hydrogen reionization fraction of 0.5.  The circles near the array name indicate the relative sizes of the element.  Note that MWA and LOFAR are cross-hatched to indicated phased array tiles.

@@

Note that in addition to the 19 elements in South Africa (and an initial prototype near Berkeley, CA), a pair of elements at the Green Bank Observatory in West Virginia have been used extensively in the validation testing, discussed below.  Also, three elements are currently under construction at the Mullard Radio Astronomy Observatory outside Cambridge, UK for additional feed and element tests.

@@

The large collecting area of a
HERA element yields nearly 5 times the sensitivity of an MWA tile and more
than 20 times that of a PAPER element, but it does so without substantially
degrading our ability to isolate and remove foreground emission on the basis of
spectral smoothness (Fig. \ref{fig:reflectometry}, right panel).  As shown in
\Mycitet{parsons_et_al2012b}, the amplitude and timescale of signal reflections
relates directly to the leakage of smooth-spectrum foregrounds into regions of Fourier space 
used to measure reionization, as discuss in Sec. \ref{sec:wedge}.
To facilitate foreground filtering, HERA's antenna element is designed to suppress reflections at long time delays.

%%%%%%%%%%FIXUP HERE
\begin{figure}[h!]
	\centering
    \includegraphics[width=0.49\textwidth]{plots/s11_compare_msip.pdf}
	%\includegraphics[width=0.49\textwidth]{plots/compare_kernels_msip_subband.pdf}
\includegraphics[width=0.5\textwidth]{plots/ps1d_with_delay_kernel.pdf}
	\vspace{-25pt}
	\caption{Left: Electromagnetic simulations (black; \citealt{ewallwice_et_al2016}) of the response of the HERA dish as a function of time delay, $\tau$,
       agree with field measurements (grey; \citealt{patra_et_al2016}). 
Right: Modeled foreground power spectra with (green) and without (blue) the antenna chromaticism over a 10\,MHz subband centered at $z=7.5$. The spectral smoothness of the dish enables foreground isolation below the level of a fiducial reionization signal (black) for most $k_{\parallel}$ modes, yielding percent-level constraints on model parameters.}
	\label{fig:reflectometry}
	\vspace{-10pt}
\end{figure}

As shown in Fig.~\ref{fig:reflectometry}, HERA's element meets this requirement in
reflectometry tests, performing comparably to the PAPER element.  The
relatively short (4.5\,m) focal height of the parabolic dish 
minimizes the timescale of feed-dish reflections while maintaining the feed's
illumination of the dish.  HERA's feed design is based on the PAPER sleeved
dipole element, inverted above a backplane, with a skirting cylinder to
improve symmetry in the polarized beam response patterns.  
%%%%%%%%%%%

Figure
\ref{fig:orbcommexptandbeammap} shows the beam pattern measured at 137 MHz with a beam mapping
system using the ORBCOMM satellite network \Mycitep{neben_et_al2016}.  Results
indicate an effective per-element collecting area of 93\,m$^2$, compared with the
theoretical maximum Airy response of 155 m$^2$.  The measured primary beam is
consistent with simulations at the 0.1\% to 0.5\% level, with a full width at
half maximum of $\sim$$10^\circ$, and a first sidelobe at -20 dB \citep{ewallwice_et_al2016,neben_et_al2016,patra_et_al2016,thyagarajan_et_al2016}.
Feed-to-feed coupling between adjacent antennas has been measured to be below -50 dB, indicating
that mutual coupling will not be problematic.
%%%%%%%%%%%%
SHOW CALCULATED BEAM PATTERNS
\begin{figure}[h!]
	\begin{tabular}{ll}
	\begin{minipage}{4.2in}
\includegraphics[width=2.1in]{plots/ref_dipole_and_hera_dish.jpg}
\includegraphics[width=2.1in]{plots/orbcomm_dish_beam_map_530cm_feed.pdf}
	\end{minipage} & 
	\begin{minipage}{2.05in}
	\caption{Left: The first of three prototype dishes at NRAO--Green Bank, used for measuring beam frequency structure with reflectometry and the beam pattern at 137\,MHz by comparing satellite signals to the reference dipole in the foreground \citep{neben_et_al2016}. Right: The measured EW power pattern plotted with dashed lines marking zenith angles of 20$^\circ$, 40$^\circ$, 60$^\circ$, 80$^\circ$.} 
	\label{fig:orbcommexptandbeammap}
	\end{minipage}
	\end{tabular}
	\vspace{-15pt}
\end{figure}

\subsubsection{Element Construction}
To minimize cost and aid in construction in a remote location, the construction materials are readily available standard construction materials, making for a very cost-effective design.  The feed is supported from three utility poles using line and hardware primarily from boating activities which can handle the load and external environment.  Given the close-packed design, each pole (except for the perimeter) are shared between three antennas.  This both reduces the number of poles, but also makes for a balanced load.  Note that the perimeter poles can be stayed with guy lines if needed.   The rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.  Two of the three intermediate posts are shared with adjacent antennas.

The center of the antenna is a cast-in-place concrete donut-shaped hub with 12 sleeves for PVC spars which provide the support for the surface and 12 horizontal spars which allow for additional support.  The twelve surface spars are made of 60mm PVC pipe, which are stressed along three support points to approximate a parabolic shape.  Note that an ideal moment-loaded beam actually attains a true parabola.  The three support points (one at the hub, one about 3 meters out radially and one at the rim-end) are at the correct height and angle for the underlying parabola and the PVC essentially acts as a spatial filter.

The shape between the spars is actually a parabola rather than a tparaboloid, so the dish surface may more properly be called a ``faceted parabola''.  If only twelve spars were used for the entire surface, simple Ruze losses would be quite high (about 10\% at 150 MHz).  Therefore, at about 2.4 meters another horizontal member is placed, which launches another parabolic spar.  The simple Ruze loss is then less than 1\%, which is much more acceptable.  Fig \ref{fig:elementplan} shows this surface scheme and the panelization (left) and the offsets (in mm) of that scheme from an ideal parabola.  The panel that is ``left out'' is the location of a door to allow access into the hub (via a small removable door and bridge).  The additional intermediate spars also make the panel sizes much more manageable and better secured.  The panel arrangement is also set by a standard width of mesh roll of 1.22m.

\begin{figure}[h!]
\centerline{
\includegraphics[height=6cm]{plots/antPlan.pdf}
\includegraphics[height=6cm]{plots/offset.pdf} 
}
\caption{\small Element panel plan (left) and residual from a true paraboloid (right).
\label{fig:elementplan}}
\end{figure}

\subsubsection{Feed}
The feed is to be deployed along two separate phases.  Initially, the feed reuses the original PAPER sleeved dipole, however in a new optimized cylindrical feed-cage configuration.  The optimization was based on main beam efficiency, cross-coupling (integrated feed pattern on the dish), standing waves (blockage size), and frequency response.  The design parameters were size of backplane, height of mast, and height of cylinder.  A range of conical structures were looked at but were ruled out since it was found they introduce additional frequency structure with no real improvement in performance.  The phase 1 design documented in a memo \cite{feedmemo} is currently under test on the 19 elements in South Africa. Additional studies are on-going at NRAO and Cambridge University.

@@

The low noise amplifier/balun integrates directly to the base of the feed dipole in the same configuration as per PAPER.  The 150m length of 75 $\Omega$ cable used in the initial design also runs down centrally and then over to the post-amplifier module, which is housed near the RFI-shielded container containing the digital electronics.  This will be replaced by the 35m cable/node design, to be discussed.

@@

As an experiment, HERA has optimized its configuration for the delay-spectrum approach to measure the EoR power spectrum.  This means that it should be on short redundant baselines.  Measurements conducted with PAPER showed that baselines of about 15m allow \cite{dillon_parsons2016}.

HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments 
to cover the $uv$-plane with sub-element sampling density (see Fig.~\ref{fig:arrayConfig}). The dense core maximizes 
sensitivity on the short baselines for which PAPER's proven foreground filtering strategy works best.
The core is supplemented by 30 additional outrigger elements 
to tile the $uv$-plane with instantaneously complete sub-aperture 
sampling out to 250$\lambda$ and complete aperture-scale sampling out to 350$\lambda$ (at 150\,MHz). This sampling 
strategy suppresses grating lobes in the synthesized beam and provides information for calibrating and correcting 
direction-dependent antenna responses \citep{dillon_parsons2016}.
Even with the sub-aperture dithering and long baselines, all 350 HERA elements can be robustly calibrated by taking advantage of 
its highly redundant configuration \citep{liu_et_al2010,zheng_et_al2014}. 
Resulting calibration errors range from $\sim5\%$ (in the core) to $\sim10\%$ (for outriggers)
of the residual fractional noise per antenna after averaging.
